---
title: "p8105_hw3_bh2852"
author: "Beicheng Huang"
date: "2023-10-14"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


##Q2

```{r Load data and cleaning}
library(p8105.datasets)
data("brfss_smart2010")

brfss = brfss_smart2010 |>
  janitor::clean_names() |> 
  rename("State" = locationabbr) |>
  rename("County" = locationdesc) |>
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), topic == "Overall Health") |> 
  mutate(response = factor(response, 
                           levels = c( "Poor", "Fair","Good", "Very good", "Excellent", ordered = TRUE)))
```


In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss |> 
  filter(year == 2002) |> 
  group_by(State) |> 
  summarise(location_N = n_distinct(County)) |> 
  filter(location_N >= 7)

brfss |> 
  filter(year == 2010) |> 
  group_by(State) |> 
  summarise(location_N = n_distinct(County)) |> 
  filter(location_N >= 7)

```
In 2002, CT, FL, MA, NC, NJ, PA were observed in 7 or more locations. In 2010, CA, CO, FL, MA, MD,NC,NE,NJ,NY,OH were observed in 7 or more locations.


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.
```{r}
excellent_only = 
  brfss |> 
  filter(response == "Excellent") |>
  group_by(year, State) |> 
  summarise(data_value_avg = mean(data_value))

# plot:
excellent_only |> 
  ggplot(aes(x = year, y = data_value_avg, color = State)) + 
  geom_line(alpha = .5) +
  geom_point() + 
  labs(x = "Year", 
       y = "Average Data Values of specific states", 
       title = "Average data values across years for each state") + 
  theme(legend.position = "bottom",
        legend.box.background = element_rect())
```
The new dataset called "Excellent_only" contains 3 variables (year, State, and data_value_avg) and 443 observations in total.

The "plot" shows the distribution of average data values of states across the years.


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r boxplot}
brfss |> 
  filter(year %in% c(2006, 2010), State == "NY") |> 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_boxplot() + 
  facet_wrap(~year) + 
  labs(x = "Response", 
       y = "Data Value",
       title = "Distribution of responses among locations in NY State (2006 & 2010)") + 
  theme(legend.position = "none",
        panel.spacing = unit(1, "lines"))
```
Both 2006 and 2010 data distribution across states show a relative normal distribution with slightly right skewed distribution towards Excellentce. For both plots, Poor responses has the smallest data value while in 2006 Good responses has the largest data value and in 2o10 the very good responses has the most data values.


##Q3

```{r Load data}
nhanes_covar = 
  read_csv(file = "nhanes_covar.csv",
           skip = 4) |> 
  janitor::clean_names() |> 
  filter(age >= 21) |> 
  drop_na()

nhanes_accel = 
  read_csv(file = "nhanes_accel.csv") |> 
  janitor::clean_names()
  
```
```{r Merge dataset}
nhanes_merge = 
  inner_join(nhanes_covar, nhanes_accel, by = "seqn")

```

```{r Factor variables and change Variable}
nhanes_merge$sex = factor(nhanes_merge$sex, 
                    levels = c("1", "2"),
                    labels = c("Male", "Female"))

nhanes_merge$education = factor(nhanes_merge$education, 
                          levels = c("1", "2", "3"),
                          labels = c("Less than High School", "High School Equivalent", "More than High School"))

nhanes_merge$seqn = as.integer(nhanes_merge$seqn)
nhanes_merge$age = as.integer(nhanes_merge$age)
```


```{r Education table for men and women}
nhanes_merge |> 
  group_by(sex, education) |> 
  summarize(count = n()) |> 
  pivot_wider(names_from = education, values_from = count) |> 
  knitr::kable(digits = 1)
```

For Less than high school, there are 27 male and 28 female. High school equivalent education has 35 male and 23 female. For more than high school education, there are 56 male and 59 female.

```{r Visualization of male and female age distribution in each education category}
nhanes_merge |> 
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot(alpha = 0.50) +
  labs(x = "Education level",
       y = "Age",
       title = "Male and Female Age Distribution in Each Education Category")
```

According to the boxplot, for education "Less than high school", the age distribution for male and female are roughly the same. For education level "High school equivalent", the mean age distribution for male is younger than that of female. For education level "More than high school", the age distribution for male and female are roughly the same with female having a slightly younger mean age.


```{r Plotting Total activity in a day for male and female in different education levels using new variable}
# create new variable
nhanes_total <-
  nhanes_merge |> 
    mutate(total_activity = rowSums(across(min1:min1440), na.rm = TRUE))

# make plots
nhanes_total |> 
ggplot(aes(x = age, y = total_activity, color = sex)) + 
  geom_point(alpha = 0.5, size = 1.5) + 
  geom_smooth(method = lm, se = FALSE) + 
  facet_wrap(~education) + 
  labs(x = "Age",
       y = "Total Activity each day (min)",
       title = "Total Minutes of Activity in a Day for Different Education Level (Male vs Female)") +
  theme(panel.spacing = unit(1, "lines"))
```
According to the plot, no matter what education you are, the total activity each day is reduced as age increased for men and women. Moreover, for high school equivalent education and higher, female have longer total activity time than men across age.


```{r Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex}
nhanes_final <-
  nhanes_merge |> 
  pivot_longer(cols = starts_with("min"), 
               names_to = "minute",
               values_to = "activity") |> 
  mutate(minute = as.integer(gsub("min", "", minute)))

nhanes_final |> 
  ggplot(aes(x = minute, y = activity, color = sex)) +
  geom_line(alpha = 0.5) + 
  geom_smooth(aes(group = sex), se = FALSE) + 
  facet_wrap(~education) + 
  labs(x = "Min",
        y = "Activity (min)",
       title = "24-hour Activity by Education Level for Male and Female")

```

According to the 24Hour activity graph, we can observe that the general activity trends in minutes are similar no matter education level and sex. All graphs show that there's lowest activity level around 250 minutes past 0h (midnight), and apart from the "less than high education" group, the other two educuation level demonstrate a higher activity trend for female in general.
